---
title: March 14, 2017 Pre-Class
author: Wanting Zhai
date: '2017-03-13'
slug: march-14-2017-pre-class
---

#Project Goals:


With this project we will simulate a famoues probability problem. This will not require knowledge of probability or statistics but only the logic to follow the steps in order to simulate this problem. This is one way to solve problems by using the computer. 

 1. **Gambler's Ruin**: Suppose you have a bankroll of $1000 and make bets of $100 on a fair game. By simulating the outcome directly for at most 5000 iterations of the game (or hands), estimate:
    a. the probability that you have "busted" (lost all your money) by the time you have placed your one hundredth bet. 
    b. the probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly. 
    c. the mean time you go bust, given that you go bust within the first 5000 hands.
    d. the mean and variance of your bankroll after 100 hands (including busts).
    e. the mean and variance of your bankroll after 500 hands (including busts).

```{r}
set.seed(1)
GR=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(bankroll < bet){break}
  }
  duration = k
  return(duration)
}

durations = numeric()
s = 5000
for(i in 1:s)
{
  nextd = GR(100,900,0.5)
  durations = c(durations,nextd)
}
```

###a
```{r}
a=length(durations)
b=length(which(durations<=100))
prob=b/a
prob
```

The probability that you have "busted" (lost all your money) by the time you have placed your one hundredth is 0.1198

###b
```{r}
a=length(durations)
b=length(which(durations<=500))
prob=b/a
prob
```

The probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly is 0.4774

###c
```{r}
a=durations[which(durations<=5000)]
mean=mean(a)
mean
```

The mean time you go bust, given that you go bust within the first 5000 hands is 757.8801

###d
```{r}
set.seed(1)
GR1=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(k ==100){break}
    if(bankroll<bet){break}
  }
  duration = k
  return(bankroll)
  
}

bankroll=numeric()
s = 5000
for(i in 1:s)
{
  nextb = GR1(100,900,0.5)
  bankroll=c(bankroll,nextb)
}

mean(bankroll)
var(bankroll)
```

The mean of your bankroll after 100 hands (including busts) is 879.2014
The variance of your bankroll after 100 hands (including busts) is 2716298


###e
```{r}
set.seed(1)
GR2=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(k ==500){break}
    if(bankroll< bet){break}
  }
  duration = k
  return(bankroll)
  
}

bankroll=numeric()
s = 5000
for(i in 1:s)
{
  nextb = GR2(100,900,0.5)
  bankroll=c(bankroll,nextb)
}

mean(bankroll)
var(bankroll)
```

The mean of your bankroll after 500 hands (including busts) is 536.1681
The variance of your bankroll after 500 hands (including busts) is 19437196

 
Note: you *must* stop playing if your player has gone bust. How will you handle this in the `for` loop?

2. Repeat the previous problem with betting on black in American roulette, where the probability of winning on any spin is 18/38 for an even payout.

```{r}
set.seed(1)
GR=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(bankroll < bet){break}
  }
  duration = k
  return(duration)
}

durations = numeric()
s = 5000
for(i in 1:s)
{
  nextd = GR(100,900,18/38)
  durations = c(durations,nextd)
}
```

###a
```{r}
a=length(durations)
b=length(which(durations<=100))
prob=b/a
prob
```

The probability that you have "busted" (lost all your money) by the time you have placed your one hundredth is 0.1744

###b
```{r}
a=length(durations)
b=length(which(durations<=500))
prob=b/a
prob
```

The probability that you have busted by the time you have placed your five hundredth bet by simulating the outcome directly is 0.6298

###c
```{r}
a=durations[which(durations<=5000)]
mean=mean(a)
mean
```

The mean time you go bust, given that you go bust within the first 5000 hands is 504.1114

###d
```{r}
set.seed(1)
GR1=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(k ==100){break}
    if(bankroll< bet){break}
  }
  duration = k
  return(bankroll)
  
}

bankroll=numeric()
s = 5000
for(i in 1:s)
{
  nextb = GR1(100,900,18/38)
  bankroll=c(bankroll,nextb)
}

mean(bankroll)
var(bankroll)
```

The mean of your bankroll after 100 hands (including busts) is 463.9543
The variance of your bankroll after 100 hands (including busts) is 864711.5


###e
```{r}
set.seed(1)
GR2=function(bet,bankroll,p)
{
  k=0
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bet + bankroll
      bet = (bet + bankroll)*0.1
    }else{
      bankroll = bankroll-bet
      bet = bet
      }
    k=k+1
    if(k ==500){break}
    if(bankroll< bet){break}
  }
  duration = k
  return(bankroll)
  
}

bankroll=numeric()
s = 5000
for(i in 1:s)
{
  nextb = GR2(100,900,18/38)
  bankroll=c(bankroll,nextb)
}

mean(bankroll)
var(bankroll)
```

The mean of your bankroll after 500 hands (including busts) is 31.46429
The variance of your bankroll after 500 hands (including busts) is 58251.21



3. **Markov Chains**. Suppose you have a game where the probability of winning on your first hand is 48%; each time you win, that probability goes up by one percentage point for the next game (to a maximum of 100%, where it must stay), and each time you lose, it goes back down to 48%. Assume you cannot go bust and that the size of your wager is a constant $100.
    a. Is this a fair game? Simulate one hundred thousand sequential hands to determine the size of your return. Then repeat this simulation 99 more times to get a range of values to calculate the expectation.
    b. Repeat this process but change the starting probability to a new value within 2% either way. Get the expected return after 100 repetitions. Keep exploring until you have a return value that is as fair as you can make it. Can you do this automatically?
    c. Repeat again, keeping the initial probability at 48%, but this time change the probability increment to a value different from 1%. Get the expected return after 100 repetitions. Keep changing this value until you have a return value that is as fair as you can make it. 

###a
```{r}
MC=function(bet,bankroll,p)
{
  k=0
  prob<-p
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bankroll +bet
      p = p+0.01
    }else{
      bankroll = bankroll-bet
      p = prob
    }
    k=k+1
    if(k==100000){break}
    if(bankroll < bet){break}
  }
  return(bankroll)
}

size = numeric()
s = 100
for(i in 1:s)
{
  nextb = MC(100,900,0.48)
  size = c(size,nextb)
}
mean(size)
```

This is not a fair game, since mean of size is 0.

###b
```{r}
set.seed(1)
proportion=numeric()
for (j in 1:20){
  size=numeric()
  s = 100
for(i in 1:s)
{
  nextb = MC(100,900,0.46+0.002*j)
  size = c(size,nextb)
}
  nextproportion=length(which(size==0))/100
  proportion=c(proportion,nextproportion)
}
proportion
```


When starting probablity is 0.05, the result returns as fair as I can.

###c
```{r}
set.seed(1)
MC1=function(bet,bankroll,p,increment)
{
  k=0
  prob<-p
  while(bet>0) {
    ProbVal=runif(1)
    if(ProbVal <= p)
    {
      bankroll = bankroll +bet
      p = p+increment
    }else{
      bankroll = bankroll-bet
      p = prob
    }
    k=k+1
    if(k==100000){break}
    if(bankroll < bet){break}
  }
  return(bankroll)
}

proportion = numeric()
for (z in 1:20){
size=numeric()
s = 100
for(i in 1:s)
{
  nextb = MC1(100,900,0.48,0.002*z)
  size = c(size,nextb)
}
nextproportion=length(which(size==0))/100
proportion=c(proportion,nextproportion)
}
proportion
```

When probability increment to 0.034, the prportion of size of return equals to 0.5, which returns as fair as I can


4. For the last two examples in the previous question, you calculated a mean value. Because you saved these final results in a vector, use the bootstrap to estimate the variance of the return in each case for your final answer. Once you have these results, which game has the smaller variance in returns?

###4
From 3(b)
```{r}
size=numeric()
s = 100
for(i in 1:s)
{
  nextb = MC(100,900,0.5)
  size = c(size,nextb)
}
set.seed(1)
b=sample(size,1000,replace=T)
mean(b)
var(b)
```

From 3(c)
```{r}
size=numeric()
s = 100
for(i in 1:s)
{
  nextb = MC1(100,900,0.48,0.034)
  size = c(size,nextb)
}
set.seed(1)
b1=sample(size,1000,replace=T)
mean(b1)
var(b1)
```

Game 3(b) has the smaller variance in returns.